---
title: "Diagrams"
author: "Rele, Chinmay"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Results

Creating a Stacked barplot:
More condensed
```{r}
# creating stacked barplot
library("tibble")
library(ggplot2)
temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
fly_data$species = factor(fly_data$species,levels = c( "dvir", "dmoj", "dwil", "dper", "dpse", "dana", "dbip", "dbia", "dtri", "deug", "dmel", "dsec", "dsim", "dmau", "dere", "dyak" ))
ggplot(fly_data, aes(fill=class, y=percent, x=species)) + geom_bar( stat="identity", position="fill") + labs( x = "Species", y = "Percent composition of Repeat",
          title = "Percent composition of each TE class within respective Nanopore assembly" )

ggsave( "accounted_unknowns_small.pdf", width=9, height=7 )


```


Creating a Stacked barplot:
### More condensed -- RAW BASE-PAIRS

```{r}
# creating stacked barplot
library("tibble")
library(ggplot2)

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
fly_data$species = factor(fly_data$species,levels = c( 'dvir', 'dmoj', 'dwil', 'dper', 'dpse', 'dana', 'dbip', 'dbia', 'dtri', 'deug', 'dmel', 'dsec', 'dsim', 'dmau', 'dere', 'dyak' ))

ggplot( fly_data, aes(fill=class, y=bp, x=species) ) + geom_bar(stat = "identity") + labs( y = "Basepairs of Repeat", x = "Species", title = "Basepair composition of each TE class within respective Nanopore assembly") + theme_minimal(base_size = 15) # + theme( axis.title.y=element_blank(),  axis.ticks.y=element_blank(), axis.text.y=element_blank(), legend.position = c(0.8, 0.2) , legend.background = element_rect(fill="white",  size=0.5, linetype="solid"))  + coord_flip()

ggsave( "accounted_unknowns_raw_bp.pdf", width =10, height = 12)
```


### Genome Size and Relative Abundance of Repeats

```{r}
library("tibble")
library(ggplot2)
library("dplyr")

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

cor.test(x=simple_data$assembly, y=simple_data$rmod_bp, method = 'spearman')


simple_data %>%
    mutate( TE_percent = rmod_bp/assembly ) %>%
    select( Species, rmod_bp, TE_percent )

ggplot( simple_data, aes( x = assembly,
                              y = rmod_bp,
                              color = Species ) ) %>%
    + theme(text = element_text(size=18) ) %>%
    + geom_point( size = 6 ) %>%
    + labs( x = "Assembly size (linear)",
                         y = "TE BasePairs (linear)",
                         color = "Species",
                         caption = "Spearman's Correlation= 0.9088235")

ggsave( "assembly_vs_reps.pdf", width=12, height=7 )


```

### Assembly size vs. all sequences (except Sat/SR) (including unknowns)

```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

temp0 = merge( fly_data, simple_data, by=c("short") )
temp0
temp0 = temp0[!(temp0$class=="Satellite/Simple_Repeat"),]
temp0

keeps = c( "short", "assembly", "bp" )
keeps
total = temp0[ keeps ]
total
total$short = factor(total$short,levels = c( "dvir", "dmoj", "dwil", "dper", "dpse", "dana", "dbip", "dbia", "dtri", "deug", "dmel", "dsec", "dsim", "dmau", "dere", "dyak" ))
total

temp1 = group_by(total, short)
temp1

temp2 = aggregate(temp1$bp, by=list(short=temp1$short, assembly=temp1$assembly), FUN=sum)
temp2

# summed = summarise( temp1,  )
# summed

ggplot( temp2, aes( x = assembly,
                              y = x,
                              color = short ) ) + theme(text = element_text(size=18) ) + geom_point(size =  6) + labs( x = "Assembly size (linear)",
                         y = "All sequences (except Sat/SR)(including unknowns)",
                         color = "Species",
                         caption = "Spearman's Correlation=0.5029412")

cor.test(x=temp2$x, y=temp2$assembly, method = 'spearman')


ggsave( "assembly_vs_all_seq(except_SatSR).pdf", width=9, height=7 )

```

### Assembly size vs. Just Sat/SR

```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

temp0 = merge( fly_data, simple_data, by=c("short") )
temp0
temp0 = temp0[(temp0$class=="Satellite/Simple_Repeat"),]
temp0

keeps = c( "short", "assembly", "bp" )
keeps
total = temp0[ keeps ]
total
total$short = factor(total$short,levels = c( "dvir", "dmoj", "dwil", "dper", "dpse", "dana", "dbip", "dbia", "dtri", "deug", "dmel", "dsec", "dsim", "dmau", "dere", "dyak" ))
total

temp1 = group_by(total, short)
temp1

temp2 = aggregate(temp1$bp, by=list(short=temp1$short, assembly=temp1$assembly), FUN=sum)
temp2

# summed = summarise( temp1,  )
# summed

cor.test(x=temp2$x, y=temp2$assembly, method = 'spearman')

ggplot( temp2, aes( x = assembly,
                              y = x,
                              color = short ) ) + theme(text = element_text(size=18) ) + geom_point(size =  6) + labs( x = "Assembly size (linear)",
                         y = "Only Satellites/Simple_Repeats",
                         color = "Species",
                         caption = "Spearman's Correlation=0.5764706",
                         title = "Assembly size vs. Just Sat/SR")



ggsave( "assembly_vs_SatSR.pdf", width=9, height=7 )

```

### Repeats vs. Genome Assembly

```{r}
library("tibble")
library(ggplot2)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

simple_data$sum = simple_data$trf + simple_data$rmod_bp

cor.test(x=simple_data$assembly, y=simple_data$sum , method = 'spearman')

ggplot( simple_data, aes( x = assembly, y = rmod_bp + trf, color = Species ) ) + theme(text = element_text(size=18) )  + geom_point(size = 6) + labs( x = "Assembly size (linear)",  y = "All Repeats BasePairs (linear)",  caption = "Spearman's Correlation= 0.8617647")

ggsave( "assembly_size_vs_all_reps.pdf", width = 9, height = 7 )

```

### Simple Repeats vs. TEs

```{r}
library("tibble")
library(ggplot2)
temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

simple_data$sum = simple_data$trf + simple_data$rmod_bp

corr = cor.test(x=simple_data$trf, y=simple_data$rmod_bp , method = 'spearman')
corr

ggplot( simple_data, aes( x = trf,
                              y = rmod_bp,
                              color = Species ) ) + theme(text = element_text(size=18) ) + geom_point(size = 6) + labs( x = "Simple Repeats BasePairs (linear)",
                         y = "TEs BasePairs (linear)",
                         caption = "Spearman's Correlation = 0.3411765")

ggsave( "TE_vs_SR.pdf", width=9, height=7 )

```

### (BARPLOT) percent of total unknown sequences of each species of all TE
```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

keeps = c( "short", "class", "bp" )
total = fly_data[ keeps ]
total$short = factor(total$short,levels = c( "dvir", "dmoj", "dwil", "dper", "dpse", "dana", "dbip", "dbia", "dtri", "deug", "dmel", "dsec", "dsim", "dmau", "dere", "dyak" ))
total = total[ total$class == "Unknown", ]
keep = c( "short", "bp" )
total = total[ keep ]
total

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
ass_size = simple_data[ c("short", "assembly")  ]

total = total %>%
    group_by( short ) %>%
    summarize( sum = sum(bp) )

merged = merge( total , ass_size , by="short")
merged

order = c( "dvir", "dmoj", "dwil", "dper", "dpse", "dana", "dbip", "dbia", "dtri", "deug", "dmel", "dsec", "dsim", "dmau", "dere", "dyak" )
unknown_percent_assembly = merged %>%
    slice( match ( order, short ) ) %>%
    mutate( percent_of_assembly = sum/assembly )
unknown_percent_assembly

unknown_percent_assembly$short = factor( unknown_percent_assembly$short, levels = unknown_percent_assembly$short[order(desc(unknown_percent_assembly$percent_of_assembly))])
unknown_percent_assembly

ggplot(data=unknown_percent_assembly, aes(x=short, y=percent_of_assembly)) + geom_bar(stat="identity", fill="steelblue") + theme_minimal() + labs( x = "Species",
                         y = "Percent of Genome Size") + coord_flip()

ggsave( "unknown_percent_assembly.pdf", width = 9, height = 12)

```

### Summaries of Assemblies
```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./assembly_summary.tab", header=TRUE, sep="\t" )
ass_sum = as_tibble(temp)
ass_sum

ggplot(data=ass_sum, aes(x=assembly_size, y=contig_count, color=avg_contig_size, size=N_50)) + geom_point() + labs( x = "Assembly Size", y="Number of Contigs", color = "Average Contig Size", size = "N50", title = "Summaries of Assemblies") + geom_label(aes(label = species))


ggsave( "assemblies_summary.pdf", width=9, height=7 )

```


### Faceted plot of correlation; facet along TE class

```{r}
library("tibble")
library(Rmisc)
library(ggplot2)
library(dplyr)
library(plyr)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

temp0 = merge( fly_data, simple_data, by=c("short") )
temp0 = temp0[!(temp0$class=="Other"),]
temp0

temp0 = aggregate( cbind(bp)~Species+class+assembly, temp0, sum )
dna = temp0 %>%
    filter( class == "DNA" )
line = temp0 %>%
    filter( class == "LINE" )
ltr = temp0 %>%
    filter( class == "LTR" )
rc = temp0 %>%
    filter( class == "RC" )
sat_sr = temp0 %>%
        filter( class == "Satellite/Simple_Repeat" )
unknown = temp0 %>%
    filter( class == "Unknown" )
chimera = temp0 %>%
    filter( class == "Chimera" )

cor.test(x=dna$assembly, y=dna$bp , method = 'spearman')
cor.test(x=line$assembly, y=line$bp , method = 'spearman')
cor.test(x=ltr$assembly, y=ltr$bp , method = 'spearman')
cor.test(x=rc$assembly, y=rc$bp , method = 'spearman')
cor.test(x=sat_sr$assembly, y=sat_sr$bp , method = 'spearman')
cor.test(x=unknown$assembly, y=unknown$bp , method = 'spearman')
cor.test(x=chimera$assembly, y=chimera$bp , method = 'spearman')

labels = c( "DNA", "LINE", "LTR", "RC", "Sat/SR", "Unknown", "Chimera" )

ggplot( temp0 , aes( x=assembly, y=bp, color = Species ) ) %>%
    + geom_point() %>%
    + geom_smooth(aes( group = class), method = "lm", se = FALSE, size = 0.9, colour = "grey", stat="smooth") %>%
    # + stat_smooth( method = 'lm' ) %>%
    + facet_wrap( ~class, scales="free", ncol = 2) %>%
    + labs( x = "Assembly_size",
            y = "Basepairs",
            color = "Species",
            title = "BP Correlation with Assembly size per Class of Repeats") # %>%
    # + geom_text( data = temp0, mapping = aes( label = labels ) )

ggsave( "faceted_class_correlation.pdf", width=9, height=7 )


```
---

### RepeatModeler Summary (faceted per class)
- How manys equences were identified as TEs?
- How many sequences were identified as Sat/SR?
- How many sequences were descriebd as Unknown?

```{r}
library("tibble")
library(ggplot2)
library(dplyr)
library(magrittr)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

more_data = merge( fly_data, simple_data, by=c("short") )
more_data = more_data[!(temp0$class=="Other"),]
more_data

grouped_fly = more_data %>%
    dplyr::group_by( short, class ) %>%
    dplyr::summarise( count = n() )

grouped_fly

temp = grouped_fly
grouped_fly = temp %>%
    filter( class != "Unknown" ) %>%
    filter( class != "Satellite/Simple_Repeat" ) %>%
    filter( class != "Other" ) %>%
    dplyr::group_by( short, class ) %>%
    dplyr::summarize( total = sum(count) )
grouped_fly

ggplot( data = grouped_fly, aes( x = class, y = total, fill = class ) ) %>%
    + geom_col() %>%
    + facet_wrap( ~short, nrow=4 ) %>%
    + labs( x = "Species",
            y = "Counts of Repeat Identified",
            fill = "Classes",
            title = "Summary of number of sequences identified per class")



ggsave( "rmodeler_summary.pdf", width=9, height=7 )

```


### Percent composition of TEs of the whole genome

```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./simple_data.tab", header=TRUE, sep="\t" )
simple_data = as_tibble( temp )
simple_data

temp = read.table( "./more_condensed.tab", header=TRUE, sep="\t" )
fly_data = as_tibble(temp)
fly_data
colnames(fly_data)[colnames(fly_data)=="species"] <- "short"
fly_data

more_data = merge( fly_data, simple_data, by=c("short") )
more_data = more_data[!(temp0$class=="Other"),]
more_data
grouped_fly = more_data %>%
    group_by( Species ) %>%
    mutate( bp_percent = bp/assembly ) %>%
    select( Species, bp_percent, class )
grouped_fly

grouped_fly %>%
    group_by( Species, class ) %>%
    dplyr::summarize( total = sum(bp_percent) )

grouped_fly = grouped_fly[!(grouped_fly$class=="Other"),]

grouped_fly$Species = factor(grouped_fly$Species,levels = c( "Drosophila_yakuba", "Drosophila_erecta", "Drosophila_mauritiana", "Drosophila_simulans", "Drosophila_sechellia", "Drosophila_melanogaster", "Drosophila_eugracilis", "Drosophila_triauraria", "Drosophila_biarmipes", "Drosophila_bipectinata", "Drosophila_ananassae", "Drosophila_pseudoobscura", "Drosophila_persimilis", "Drosophila_willistoni", "Drosophila_mojavensis", "Drosophila_virilis" ))

grouped_fly

ggplot( grouped_fly , aes( y=bp_percent, fill=class, x = Species) ) + geom_bar(stat = "identity") + labs( fill = "Classes",  y = "Percent of Genome",
          title = "Percent composition of repeats within Genome" )  + coord_flip()

ggsave( "percent_composition_repeats_in_genome.pdf", width=9, height=7 )


```


### Plotting Unknown percent and BP

```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./unknown_bp_percent.tab", header=TRUE, sep="\t" )
unknowns = as_tibble( temp )

ggplot(unknowns, aes(x = species)) +
  geom_col(aes( y = bp_mb, fill="redfill")) +
  geom_text(aes(y = bp_mb, label = bp_mb), fontface = "bold", vjust = 0.5, hjust = 1.4, color = "black", size = 7) +
  geom_line(aes(y = percent * 300, group = 1, color = 'blackline')) +
  geom_text(aes(y = percent * 300, label = round(percent, 2)), hjust = 1, color = "black", size = 6) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~ . / 300)) +
  scale_fill_manual('', labels = 'Repeats content (Mb)', values = "#b3bcff") +
  scale_color_manual('', labels = 'Percent of Genome', values = 'black') +
  theme_minimal(base_size = 20) + theme( legend.position = "bottom", axis.title.y=element_blank() ) + coord_flip()

ggsave( "unknown_bp_percentp.pdf", width = 9, height = 12)

```


### BARPLOT -- Lengths of each class

```{r}
library("tibble")
library(ggplot2)
library(dplyr)

temp = read.table( "./lengths.tab", header=TRUE, sep="\t" )
lens = as_tibble( temp )

lens

## remove unneeded classes
lens = lens[ !( lens$class == "buffer" ), ]
lens = lens[ !( lens$class == "DNA?" ), ]
lens = lens[ !( lens$class == "Other" ), ]
lens = lens[ !( lens$class == "rRNA" ), ]
lens = lens[ !( lens$class == "SINE" ), ]
lens = lens[ !( lens$class == "SINE?" ), ]

lens



ggplot( lens , aes(x=class, y=length)) +
    geom_boxplot() +
    scale_y_log10() +
    ggtitle( "Lenght of each Class of Putative TE", subtitle = "in log_10" )    




ggsave( "class_lengths.pdf", width = 9, height = 6)

```
